This script displays the [*Maroon*](https://chicagomaroon.com/)'s R code we used in the process of cleaning and processing data for [our project](https://chicagomaroon.github.io/data-visualizations/2025/property-scroller/) on the University's property footprint over time.

The dataset we compiled is available [here](https://drive.google.com/file/d/1LZ-jASBVvhj4Y5aN7GEAkoQvPRumb9Uo/view?usp=sharing) in GeoJSON format. This document references many other files (mostly spatial files for individual buildings we had to draw by hand), which we are happy to provide upon request. Please reach out to data\@chicagomaroon.com with any questions, concerns, or requests for further information.

```{r message=FALSE, warning=FALSE, results='hide'}
library(tidyverse)
library(sf)
library(osmdata)
library(tmap)

wd <- "[YOUR WORKING DIRECTORY]"

# Get relevant bounding box
community_areas <- st_read(paste0(wd, "/Sources/chicago_community_areas.geojson"))

neighborhood_bbox <- community_areas %>% 
  filter(community %in% c("GRAND BOULEVARD",
                          "HYDE PARK", 
                          "KENWOOD", 
                          "OAKLAND",
                          "WASHINGTON PARK", 
                          "WOODLAWN")) %>%
  st_union() %>%
  st_bbox()

# Note: "neighborhood" is a very large dataframe
neighborhood <- opq(bbox = neighborhood_bbox) %>%
  add_osm_feature(key = 'building') %>%
  osmdata_sf()

neighborhood_polygons <- neighborhood$osm_polygons %>% 
  st_transform(st_crs(4326))
neighborhood_multipolygons <- neighborhood$osm_multipolygons %>% 
  st_transform(st_crs(4326))
rm(neighborhood)
# Note: the "osm_points" field doesn't seem to have any important info

# Try converting polygons to multipolygons so we can combine everything into one table
neighborhood_buildings <- neighborhood_polygons %>% st_cast("MULTIPOLYGON") %>% bind_rows(neighborhood_multipolygons)

ggplot() + 
  geom_sf(data = neighborhood_bbox %>% st_as_sfc()) + 
  geom_sf(data = neighborhood_buildings)

neighborhood_parking_lots <- opq(bbox = neighborhood_bbox) %>%
  add_osm_feature(key = 'amenity', value = 'parking') %>%
  osmdata_sf()

neighborhood_parking_lots_polygons <- neighborhood_parking_lots$osm_polygons %>% 
  st_transform(st_crs(4326))
neighborhood_parking_lots_multipolygons <- neighborhood_parking_lots$osm_multipolygons %>% 
  st_transform(st_crs(4326))

# Add non-building/parking geometries by OSM ID
query <- opq_osm_id(type = "way", id = c(
  "140438500", # Lab School Field
  "153090082", # Campus South Athletic Field
  "166212110" # Field at Woodlawn Charter Campus
)) %>% opq_string() %>% osmdata_sf()
outliers <- query$osm_polygons %>% st_cast("MULTIPOLYGON") %>%
  mutate(name = case_when(name == "Campus South Atheletic Field" ~ "Campus South Athletic Field",
                          .default = name))
rm(query)

# Create full buildings/parking lot crosswalk within neighborhood bbox

all_geometries <- bind_rows(neighborhood_buildings,
                       neighborhood_parking_lots_polygons %>% st_cast("MULTIPOLYGON"),
                       neighborhood_parking_lots_multipolygons,
                       outliers) %>% 
  mutate(id = row_number(), osm_id = as.numeric(osm_id))

unique_ids <- all_geometries %>% st_drop_geometry() %>% 
  select(id, osm_id) %>%
  distinct(osm_id, .keep_all = TRUE) %>% 
  select(id) %>% 
  pull()

# Make a crosswalk of all relevant geometries in the target area
xwalk <- all_geometries %>% filter(id %in% unique_ids)

# Remove unnecessary variables to save space
rm(neighborhood, 
   neighborhood_polygons, 
   neighborhood_multipolygons, 
   neighborhood_buildings, 
   neighborhood_parking_lots_polygons, 
   neighborhood_parking_lots_multipolygons, 
   outliers,
   all_geometries, 
   unique_ids)
```

```{r message=FALSE, warning=FALSE}
# Plot the crosswalk
ggplot() + 
  geom_sf(data = neighborhood_bbox %>% st_as_sfc()) + 
  geom_sf(data = xwalk)
```

Now that we have a single table containing all of the relevant geometries in nearby neighborhoods, we turn to cleaning up anomalies in the OSM data (for example, Burton-Judson Hall and Cathey Dining Commons being one building), reading our hand-drawn geometries, and combining everything into one dataset.

```{r message=FALSE, warning=FALSE, results='hide'}
# Separate BJ and Cathey
bj_cathey <- xwalk %>% filter(osm_id == 1968186) %>% st_geometry() # Combined geometry
cathey_template <- st_read(paste0(wd, "template_geometries/cathey.geojson")) # Geometry containing Cathey and none of BJ

bj <- bj_cathey %>% st_difference(cathey_template) %>%
  st_as_sf() %>% 
  st_cast("MULTIPOLYGON") %>% 
  mutate(custom_id = 16, name = "Burton-Judson Courts") %>% 
  rename(geometry = x)
cathey <- bj_cathey %>% st_intersection(cathey_template) %>% 
  st_as_sf() %>%
  st_cast("MULTIPOLYGON") %>% 
  mutate(custom_id = 17, name = "Arley D. Cathey Dining Commons") %>% 
  rename(geometry = x)

# Add student wellness center
lying_in_hospital <- xwalk %>% filter(osm_id == "144132398") %>% st_geometry()
wellness_template <- st_read(paste0(wd, "template_geometries/wellness-center.geojson")) # Geometry containing wellness center and some of surrounding building

wellness <- wellness_template %>% st_difference(lying_in_hospital) %>%
  st_as_sf() %>% 
  st_cast("MULTIPOLYGON") %>% 
  mutate(custom_id = 9, name = "Student Wellness Center") %>% 
  select(custom_id, name)

# Separate Snell and Hitchcock
snitchcock <- xwalk %>% filter(osm_id == 154421807) %>% st_geometry() # Combined geometry
snell_template <- st_read(paste0(wd, "template_geometries/snell.geojson")) # Geometry containing Snell and none of Hitchcock

snell <- snitchcock %>% st_intersection(snell_template) %>%
  st_as_sf() %>% 
  st_cast("MULTIPOLYGON") %>% 
  mutate(custom_id = 18, name = "Snell Hall") %>% 
  rename(geometry = x)
hitchcock <- snitchcock %>% st_difference(snell_template) %>%
  st_as_sf() %>% 
  st_cast("MULTIPOLYGON") %>% 
  mutate(custom_id = 19, name = "Hitchcock Hall") %>% 
  rename(geometry = x)

# Separate 5439-45 and 5447-57 S Woodlawn
wl_apartments <- xwalk %>% filter(osm_id == 211225247) %>% st_geometry() # Combined geometry
wl_apartments_template <- st_read(paste0(wd, "template_geometries/5447-57-s-woodlawn.geojson")) # Geometry containing 5447-57 apartments and none of 39-45

wl_39_45 <- wl_apartments %>% st_difference(wl_apartments_template) %>%
  st_as_sf() %>% 
  st_cast("MULTIPOLYGON") %>% 
  mutate(custom_id = 39, name = "5439-45 South Woodlawn Avenue") %>% 
  rename(geometry = x)
wl_47_57 <- wl_apartments %>% st_intersection(wl_apartments_template) %>%
  st_as_sf() %>% 
  st_cast("MULTIPOLYGON") %>% 
  mutate(custom_id = 40, name = "5447-57 South Woodlawn Avenue") %>% 
  rename(geometry = x)

# Combine halves of Woodward Court
woodward_court <- st_read(paste0(wd, "custom_geometries/woodward_court.geojson")) %>% 
  st_union() %>% 
  st_as_sf() %>% 
  mutate(custom_id = 30, name = "Woodward Court") %>% 
  rename(geometry = x)

# Read hand-drawn geometries for former Stagg Field and Hyde Park Labs
stagg_hp_labs <- st_read(paste0(wd, "custom_geometries/stagg-and-hydeparklabs.geojson")) %>%
  st_transform(4326) %>%
  st_cast("MULTIPOLYGON")

# Read hand-drawn geometries and combine into one dataframe
custom <- st_read(paste0(wd, "custom_geometries/custom_buildings_7_23_24.geojson")) %>%
  st_transform(4326) %>%
  st_cast("MULTIPOLYGON") %>%
  select(custom_id, name) %>% 
  bind_rows(bj, cathey, snell, hitchcock, wl_39_45, wl_47_57, woodward_court, stagg_hp_labs) %>% 
  filter(custom_id != 9) %>% # Filter out old wellness center
  bind_rows(wellness)

# Read current buildings dataset
current <- read_csv(paste0(wd, "spreadsheet_data/uchicago_buildings_1_27_25.csv")) %>%
  filter(!is.na(osm_id) | !is.na(custom_id)) # If there's no ID, we don't have the geometry prepared for this building, so we exclude it

# Join data with geometries
property <- current %>% left_join(xwalk %>% bind_rows(custom) %>%
                                    select(osm_id, custom_id),
                                  by = c("osm_id", "custom_id")) %>%
  mutate(id = coalesce(osm_id, custom_id) %>% as.character()) %>%
  st_as_sf() %>% 
  st_make_valid()

property <- property %>% cbind(property %>% st_geometry() %>% 
                                 st_point_on_surface() %>% 
                                 st_coordinates() %>% 
                                 as_tibble() %>% 
                                 rename(c(lon = X, lat = Y)))

# Download property in GeoJSON format
st_write(property, paste0(wd, "results/property_1_27_25.geojson"))
```

```{r message=FALSE, warning=FALSE}
ggplot() + 
  geom_sf(data = neighborhood_bbox %>% st_as_sfc()) + 
  geom_sf(data = property)
```

Finally, we run various transformations on `property` to create the public dataset and the display dataset for the story.

```{r message=FALSE, warning=FALSE, results='hide'}
# Create public version of dataset ---------------------------------------------
public_dataset <- property %>% filter(maybe_not_owned_by_uni == FALSE) %>%
  mutate(year_bought = as.numeric(year_bought),
         year_built = as.numeric(year_built),
         year_destroyed = as.numeric(year_destroyed),
         year_sold = as.numeric(year_sold)) %>% 
  select(id, 
         osm_id, 
         custom_id, 
         name, 
         other_name, 
         address, 
         earliest_year_seen, 
         latest_year_seen, 
         year_bought, 
         year_built, 
         year_destroyed, 
         year_sold, 
         year_start_guess, 
         year_end_guess, 
         currently_owned, 
         currently_exists, 
         missing_year_info, 
         year_source, 
         building_source)

st_write(public_dataset, paste0(wd, "results/public_dataset_1_27_25.geojson"), append = FALSE)

# Summarize year info ----------------------------------------------------------

property_with_years <- property %>% filter(maybe_not_owned_by_uni == FALSE) %>%
  mutate(year_bought = as.numeric(year_bought),
         year_built = as.numeric(year_built),
         year_destroyed = as.numeric(year_destroyed),
         year_sold = as.numeric(year_sold)) %>% 
  mutate(year_start = case_when(!is.na(year_bought) ~ year_bought,
                                !is.na(year_built) ~ year_built,
                                !is.na(earliest_year_seen) ~ earliest_year_seen,
                                .default = NA),
         year_end = case_when(currently_owned == TRUE ~ 2025,
                              !is.na(year_sold) ~ year_sold,
                              !is.na(year_destroyed) ~ year_destroyed,
                              !is.na(latest_year_seen) ~ latest_year_seen,
                              .default = NA),
         year_start_confirmed = year_start,
         year_end_confirmed = year_end) %>%
  select(id, osm_id, custom_id, 
         name, address, 
         earliest_year_seen, latest_year_seen, 
         year_bought, year_built, year_destroyed, year_sold, 
         year_start_guess, year_end_guess, 
         year_start, year_end, year_start_confirmed, year_end_confirmed,
         currently_owned, years_confirmed, 
         quad_flag, recent_residential_sales, charter_schools_flag, commercial_buildings_flag, 
         currently_exists,
         lon, lat)

# Make display dataset ---------------------------------------------------------

property_display <- property_with_years %>% mutate(year_start = pmin(year_start, year_start_guess, na.rm = TRUE), year_end = pmax(year_end, year_end_guess, na.rm = TRUE)) %>%
  select(-c(earliest_year_seen, 
            latest_year_seen, 
            year_bought, 
            year_built, 
            year_destroyed, 
            year_sold, 
            year_start_guess, 
            year_end_guess, 
            year_start_confirmed, 
            year_end_confirmed))

st_write(property_display, paste0(wd, "results/display_dataset_1_27_25.geojson"), append = FALSE)
```
